Mamary Gland Development Exploratory Data Analysis

### 2017-01-27 presentations/LabMeeting27012016.Rmd
final version
Latest update before Labmeeting

@@

@@

>>>>

# Experimental Setup
![](img/setup0.pdf)

@@

# Experimental Setup
![](img/setup3.pdf)

@@

# Outlook
## Analysis
- Identify clusters and relate them to cell types
- Differentially expressed genes within each cell type
- Differentiation trajectories
    - e.g. in the luminal department
- Testing specific hypotheses

## Future
- Plan experiment for tumor development

#

```{r echo=FALSE, warning=FALSE}
Bcl11a <- filter(fD, symbol=="Bcl11a") %>% .$id
pD.filtered$Bcl11a <- t(m[,as.character(pD.filtered$barcode)])[,Bcl11a]/pD.filtered$sf
tsnePlot <- ggplot(pD.filtered, aes(x=tSNE1, y=tSNE2, color=log2(Bcl11a+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    scale_color_viridis() +
    theme_bw()
tsnePlot
```

### 2017-01-27 src/QCAnalysis.Rmd
Latest update before Labmeeting

# Quality Overview
## Sequencing Depth and Genes detected
```{r}
pD$UmiSums<- colSums(m)
pD$GenesDetected <- colSums(m!=0)
genesDetected <- ggplot(pD, aes(x=SampleID,y=log10(GenesDetected),fill=Condition)) +
    geom_violin(draw_quantiles=0.5)+
    theme_bw()
LibrarySize <- ggplot(pD, aes(x=SampleID,y=log10(UmiSums),fill=Condition)) +
    geom_violin(draw_quantiles=0.5)+
    theme_bw()
gdLibSize <- ggplot(pD, aes(y=log10(GenesDetected), x=log10(UmiSums), color=Condition, shape=Replicate)) +
    geom_point() +
    geom_smooth(method=loess,inherit.aes=FALSE,aes(y=log10(GenesDetected),x=log10(UmiSums))) +
    theme_bw()
genesDetected
LibrarySize
gdLibSize
```

@@

gdHist <- ggplot(pD, aes(x=log10(GenesDetected),y=..density..)) +
    geom_histogram(fill="white",color="black") +
    geom_vline(data=smryByGroup,aes(xintercept=threshold_GenesDetected),color="red",lty="longdash") +
    facet_wrap(~Condition) +
    theme_bw()
gdHist

libSizeHist <- ggplot(pD, aes(x=log10(UmiSums),y=..density..)) +
    geom_histogram(fill="white",color="black") +
    geom_vline(data=smryByGroup,aes(xintercept=threshold_UmiSums),color="red",lty="longdash") +
    facet_wrap(~Condition) +
    theme_bw()
libSizeHist

@@

fPCA.lac <- log2(t(m.lac.norm[fD.lac$highVar,])+1)
fPCA.lac <- scale(fPCA.lac,scale=TRUE,center=TRUE)
set.seed(rnd_seed)
tsn.lac <- Rtsne(fPCA.lac,perplexity=floor(nrow(fPCA.lac)/5))

@@

tsnPlot <- ggplot(pD.lac, aes(x=tSNE1, y=tSNE2, color=PassAll, shape=Replicate)) +
    geom_point(alpha=0.8,size=3) +
    scale_color_manual(values=c("dodgerblue","red")) +
    theme_bw() 
tsnPlot

@@

highVar <- dplyr::arrange(fD.filtered,desc(dm)) %>% .$id
topX <- floor(0.2*length(highVar))

fD.filtered$highVar <- fD.filtered$id %in% highVar[1:topX]
VarPlot <- ggplot(fD.filtered, aes(x=log10(meanExpression),y=log10(CV2), color=highVar)) +
    geom_point() +
    theme_bw()
VarPlot

fPCA <- log2(t(m.norm[fD.filtered$highVar,])+1)
fPCA <- scale(fPCA,scale=TRUE,center=TRUE)
pc <- prcomp(fPCA)
set.seed(rnd_seed)
tsn <- Rtsne(fPCA,perplexity=floor(nrow(fPCA)/5))

@@

pcaPlot <- ggplot(pD.filtered, aes(x=PC1, y=PC2, color=log2(Csn2+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw() 
pcaPlot

>>>>

# Quality Overview
## Sequencing Depth and Genes detected
```{r}
pD$UmiSums<- colSums(m)
pD$GenesDetected <- colSums(m!=0)
genesDetected <- ggplot(pD, aes(x=SampleID,y=GenesDetected,fill=Condition)) +
    geom_violin(draw_quantiles=0.5)+
    scale_y_log10() +
    theme_bw()
LibrarySize <- ggplot(pD, aes(x=SampleID,y=UmiSums,fill=Condition)) +
    geom_violin(draw_quantiles=0.5)+
    scale_y_log10() +
    theme_bw()
gdLibSize <- ggplot(pD, aes(y=log10(GenesDetected), x=log10(UmiSums), color=Condition, shape=Replicate)) +
    geom_point() +
    geom_smooth(method=loess,inherit.aes=FALSE,aes(y=log10(GenesDetected),x=log10(UmiSums))) +
    theme_bw()
genesDetected
LibrarySize
gdLibSize
```

@@

gdHist <- ggplot(pD, aes(x=GenesDetected,y=..density..)) +
    geom_histogram(fill="white",color="black") +
    geom_vline(data=smryByGroup,aes(xintercept=10^threshold_GenesDetected),color="red",lty="longdash") +
    scale_x_log10() +
    facet_wrap(~Condition) +
    theme_bw()
gdHist

libSizeHist <- ggplot(pD, aes(x=UmiSums,y=..density..)) +
    geom_histogram(fill="white",color="black") +
    geom_vline(data=smryByGroup,aes(xintercept=10^threshold_UmiSums),color="red",lty="longdash") +
    scale_x_log10() +
    facet_wrap(~Condition) +
    theme_bw()
libSizeHist

@@

fPCA.lac <- log2(t(m.lac.norm[fD.lac$highVar,])+1)
fPCA.lac <- scale(fPCA.lac,scale=TRUE,center=TRUE)
set.seed(rnd_seed)
tsn.lac <- Rtsne(fPCA.lac,perplexity=25)

@@

tsnPlot <- ggplot(pD.lac, aes(x=tSNE1, y=tSNE2, color=PassAll, shape=Replicate)) +
    geom_point(alpha=0.8,size=3) +
    scale_color_manual(values=c("grey50","black")) +
    theme_bw() 
tsnPlot

@@

#HVG definition by Brennecke
brennecke <- BrenneckeHVG(m.norm,fdr=0.1,minBiolDisp=0.25)
fD.filtered$highVar <- fD.filtered$id %in% brennecke

VarPlot <- ggplot(fD.filtered, aes(x=meanExpression,y=CV2, color=highVar)) +
    geom_point() +
    scale_x_log10() +
    scale_y_log10() +
    theme_bw()
VarPlot

fPCA <- log2(t(m.norm[fD.filtered$highVar,])+1)
fPCA <- scale(fPCA,scale=TRUE,center=TRUE)
pc <- prcomp(fPCA)
set.seed(rnd_seed)
tsn <- Rtsne(fPCA,perplexity=25)

@@

pcaPlot <- ggplot(pD.filtered, aes(x=PC1, y=PC3, color=log2(Csn2+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw() 
pcaPlot
